A Hybrid Ensemble Learning Approach to Star-galaxy Classification

### 2015-03-31 paper/main.tex
add magnitude cut and z cut.

\documentclass[useAMS,usenatbib]{mn2e}
\usepackage[letterpaper,margin=0.65in]{geometry}
\usepackage[ansinew]{inputenc}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{color}
\usepackage{amssymb}
\usepackage[normalem]{ulem}
\usepackage{natbib}

@@

%%\begin{table}
%%  \caption{The definition of the number of stars/galaxies
%%           classified and misclassified as stars/galaxies.}
%%  \centering
%%  \begin{tabular}{ l | c | c }
%%    \hline
%%                           & True galaxies & True stars \\ \hline
%%    Classified as galaxies & $N_g$ & $M_s$ \\ \hline
%%    Classified as stars & $M_g$ & $N_s$ \\
%%    \hline
%%  \end{tabular}
%%  \label{table:def_numbers}
%%\end{table}

@@

We use photometric data from
the Canada-France-Hawaii Telescope Lensing Survey
\cite[CFHTLenS\footnote{http://www.cfhtlens.org/};][]
{Heymans2012,Erben2013,Hildebrandt2012}.
This catalog consits of more than twenty five million objects
with a limiting magnitude of $i_{\text{AB}} \approx 25.5$. 
It covers a total of 154 square degress
in the four fields of CFHT Legacy Survey~\citep[CFHTLS;][]{gwyn2012canada}
observed in the five photometric bands:
$u$, $g$, $r$, $i$, and $z$.

@@

%% rich training data
\begin{table*}
  \caption{A summary of the classification performance metrics
           for the four individual methods
           and the four different classification combination methods
           as applied to the CFHTLenS data,
           with no cut applied to the training data set.
           The definition of the metrics is summarized in
           Table~\ref{table:metrics}.
           The bold entries highlight the best performance values
           within each column.}
  \centering
%  \begin{tabular}{l c c c c c c}
%    Classifier & AUC & MSE &
%    $p_{g}\left(c_g=0.9964\right)$ & $p_{s}\left(c_s=0.7145\right)$ &
%    $p_{g}\left(c_g=0.9600\right)$ & $p_{s}\left(c_s=0.2500\right)$ \\
%    \hline
%    TPC  & 0.9870 & 0.02081 & 0.9717 & 0.9843 & 0.9925 & 0.9977\\
%    SOMc & 0.9683 & 0.04520 & 0.9099 & 0.8487 & 0.9781 & 0.9844\\
%    HB   & 0.8954 & 0.08777 & 0.9048 & 0.7407 & 0.9593 & 0.7609\\
%    Morphology   & -      & 0.0397 & 0.9597 & 0.9666 & - & -\\
%    WA   & 0.9718 & 0.03292 & 0.9703 & 0.9876 & 0.9766 & 1.0000\\
%    BoM  & 0.9870 & 0.02081 & 0.9733 & 0.9843 & 0.9925 & 0.9977\\
%    Stacking  & \textbf{0.9878} & 0.02043 & 0.9729 & 0.9909 &
%    \textbf{0.9928} & 1.0000\\
%    BMC  & 0.9849 & \textbf{0.01911} & \textbf{0.9797} & \textbf{0.9950} &
%    0.9916 & 1.0000\\
%  \end{tabular}
  \input{tables/metrics_all.tex}
  \label{table:metrics_all}
\end{table*}

@@

The simplest of the combination techniques, WA and BoM,
generally do not perform better than \texttt{TPC}.
It is also interesting that,
even with binning the parameter space and selecting the best model
within each bin,
BoM almost always chooses \texttt{TPC} as the best model,
and therefore gives the same performance as \texttt{TPC} in the end.
However, our BMC and stacking techniques have a similar performance
and often outperform \texttt{TPC}.
We observe that our stacking algorithm shows the best performance
as measured by the AUC,
while BMC shows the best performance
as measured by the MSE.

\begin{figure*}
  \begin{minipage}[b]{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/som_colors.eps}
    \caption{A two-dimensional 10$\times$10 SOM representation
             showing the mean $i$-band magnitude (top left) and
             the mean values of three colors,
             $u-g$, $g-r$, and $i-z$,
             for the cross validation data.}
    \label{fig:som_colors}
  \end{minipage}
  \hfill
  \begin{minipage}[b]{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/weights.eps}
    \caption{A two-dimensional 10$\times$10 SOM representation
             showing the relative weights for the BMC combination technique
             applied to the four individual methods for the CFHTLenS data.}
    \label{fig:weights}
  \end{minipage}
\end{figure*}

In Figure~\ref{fig:som_colors}, we show
the mean CFHTLenS $i$-band magnitude in the top right panel,
and the mean $u-g$, $g-r$, and $i-z$ colors in the remaining three panels.
These two-dimensional maps clearly shows
the ability of the SOM to preserve
relationships between sources
when it projects the full nine-dimensional space to
the two-dimensional map.
The SOM mapping is a non-linear representation of
all magnitudes and colors,
and thus the CHFTLenS $i$-band magnitude and color maps should 
only be used to provide guidance.

We can also use the same SOM to determine the relative weights for 
the four individual classification methods for each cell.
We present the four weight maps for the BMC technique
in Figure~\ref{fig:weights}.
In these maps, a redder color indicates a higher weight,
or equivalently that the corresponding weight performs better in that region.
These weight maps demonstrate the variation in
the performance of the individual techniques across
the two-dimensional parameter space defined by the SOM.
Not surprisingly, the morphological separation
performs best in the lower right corner of the $i$-band magnitude map,
which corresponds to the brightest CFHTLenS magnitudes $i \lesssim 20$.
On the other hand, \texttt{TPC} seems to perform best
in the region that corresponds to intermediate magnitudes
$20\gtrsim i \lesssim22.5$ and $1.5<u-g<3.0$.
Our unsupervised learning method \texttt{SOMc}
performs relatively better at fainter magnitudes $i>21.5$
with $0<u-g<0.5$ and $0<g-r<0.5$.
The BMC technique assigns almost no weight to
the worst-performing algorithm, \texttt{HB},
when there is a wealth of training data.

@@

It is very costly in terms of telescope time to
obtain a large sample spectroscopic observations
down to the limiting magnitude of a photometric sample.
Thus, we investigate the impact of training set quality
by considering a more realistic case
where the training data is available
only for a small number of objects at bright magnitudes.
To emulate this scenario,
we only use objects that have spectroscopic labels from the VVDS
and impose a magnitude cut of $i < 22.0$ in the training data,
leaving us a training set with 6,361 objects.

@@

@@

@@

We thank the referee for a careful reading of the manuscript
and comments that improved this work.
We thank Nacho Sevilla \textcolor{red}{Jacobo? anyone else?} for helpful and insightful conversations.
We acknowledge support from the National Science Grant No.
\textcolor{red}{XXXX}.

>>>>

\documentclass[useAMS,usenatbib]{mn2e}
\usepackage[letterpaper,margin=0.65in]{geometry}
\usepackage[ansinew]{inputenc}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{color}
\usepackage{amssymb}
\usepackage[normalem]{ulem}
\usepackage{natbib}
\usepackage{url}

@@

@@

We use photometric data from
the Canada-France-Hawaii Telescope Lensing Survey
\cite[CFHTLenS\footnote{http://www.cfhtlens.org/};][]
{Heymans2012,Erben2013,Hildebrandt2012}.
This catalog consits of more than twenty five million objects
with a limiting magnitude of $i_{\text{AB}} \approx 25.5$. 
It covers a total of 154 square degrees
in the four fields of CFHT Legacy Survey~\citep[CFHTLS;][]{gwyn2012canada}
observed in the five photometric bands:
$u$, $g$, $r$, $i$, and $z$.

@@

%% rich training data
\begin{table*}
  \caption{A summary of the classification performance metrics
           for the four individual methods
           and the four different classification combination methods
           as applied to the CFHTLenS data,
           with no cut applied to the training data set.
           The definition of the metrics is summarized in
           Table~\ref{table:metrics}.
           The bold entries highlight the best performance values
           within each column.}
  \centering
  \input{tables/metrics_all.tex}
  \label{table:metrics_all}
\end{table*}

@@

The simplest of the combination techniques, WA and BoM,
generally do not perform better than \texttt{TPC}.
It is also interesting that,
even with binning the parameter space and selecting the best model
within each bin,
BoM almost always chooses \texttt{TPC} as the best model,
and therefore gives the same performance as \texttt{TPC} in the end.
However, our BMC and stacking techniques have a similar performance
and often outperform \texttt{TPC}.
\textcolor{blue}{
Although \texttt{TPC} shows the best performance as measured by the AUC,
BMC shows the best performance in all other metrics.
}

\begin{figure*}
  \begin{minipage}[b]{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/som_colors.eps}
    \caption{A two-dimensional 10$\times$10 SOM representation
             showing the mean $i$-band magnitude (top left),
             the fraction of true stars in each cell (top right),
             and the mean values of $u-g$ and $g-r$
             for the cross validation data.}
    \label{fig:som_colors}
  \end{minipage}
  \hfill
  \begin{minipage}[b]{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/weights.eps}
    \caption{A two-dimensional 10$\times$10 SOM representation
             showing the relative weights for the BMC combination technique
             applied to the four individual methods for the CFHTLenS data.}
    \label{fig:weights}
  \end{minipage}
\end{figure*}

In Figure~\ref{fig:som_colors}, we show
the mean CFHTLenS $i$-band magnitude in the top left panel
\textcolor{blue}{
and the fraction of stars in each cell in the top right panel.
}
The bottom two panels show the mean $u-g$ and  $g-r$ colors.
These two-dimensional maps clearly shows
the ability of the SOM to preserve relationships between sources
when it projects the full nine-dimensional space to the two-dimensional map.
The SOM mapping is a non-linear representation of all magnitudes and colors,
and thus these SOM maps should only be used to provide guidance.

We can also use the same SOM to determine the relative weights for 
the four individual classification methods for each cell.
We present the four weight maps for the BMC technique
in Figure~\ref{fig:weights}.
In these maps, a redder color indicates a higher weight,
or equivalently that the corresponding weight performs better in that region.
These weight maps demonstrate the variation in
the performance of the individual techniques across
the two-dimensional parameter space defined by the SOM.
Not surprisingly, the morphological separation
performs best in the top left corner of the $i$-band magnitude map,
which corresponds to the brightest CFHTLenS magnitudes $i \lesssim 20$.
On the other hand, \texttt{TPC} seems to perform best
in the region that corresponds to intermediate magnitudes
$20\gtrsim i \lesssim22.5$ and $1.5<u-g<3.0$.
Our unsupervised learning method \texttt{SOMc}
performs relatively better at fainter magnitudes $i>21.5$
with $0<u-g<0.5$ and $0<g-r<0.5$.
\textcolor{blue}{
Although \texttt{HB} shows the worst performance
when there exists ample training data,
BMC still utilizes information from \texttt{HB}
to produce an overall improvement in performance,
especially at intermediate magnitudes $20\gtrsim i \lesssim20$.
}

\subsection{The Quality of Training Data}
  \label{section:quality_training}

It is very costly in terms of telescope time to
obtain a large sample spectroscopic observations
down to the limiting magnitude of a photometric sample.
Thus, we investigate the impact of training set quality
by gradually decreasing the size of the training set.

\begin{figure*}
  \begin{minipage}[b]{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/perform_mag_cut.eps}
    \caption{The classification performance metrics for TPC, HB, morphology,
      and BMC as applied to the CFHTLenS data in the VVDS field
      with various magnitude cuts.
      We show only one of the four combination methods, BMC,
      which has the best overall performance.}
    \label{fig:perform_mag_cut}
  \end{minipage}
  \hfill
  \begin{minipage}[b]{0.49\linewidth}
    \includegraphics[width=\textwidth]{figures/perform_z_cut.eps}
    \caption{Same as Figure.~\ref{fig:perform_mag_cut} using
      $z_{\text{spec}}$ cuts.}
    \label{fig:perform_z_cut}
  \end{minipage}
\end{figure*}

@@

We also investigate the impact of training set quality
by analyzing in detail a more realistic case
where the training data is available
only for a small number of objects at bright magnitudes.
To emulate this scenario,
we only use objects that have spectroscopic labels
from the VVDS 0226-04 field (which is located within the CFHTLS W1 field)
and impose a magnitude cut of $i < 22.0$ in the training data,
leaving us a training set with 6,361 objects.

@@

\textcolor{blue}{begin new text}

We have presented and analyzed a novel star-galaxy classification framework
for combining classifications from the CFHTLenS project.
In particular, we use four independent star-galaxy classification techniques:
a morphological separation method;
\texttt{TPC}, a supervised machine learning technique
based on prediction trees and a random forest;
\texttt{SOMc}, an unsupervised machine leraning approach
based on self-organizing maps and a random atlas;
and \texttt{HB}, a hierarichical bayesian template-fitting method
that we have modified and parallized.
Both \texttt{TPC} and \texttt{SOMc} are currently availble within
a software package named
\texttt{MLZ}\footnote{http://lcdm.astro.illinois.edu/code/mlz.html}.
Our implementaion of \texttt{HB} and \texttt{BMC},
as well as IPython notebooks that were used to
produce the results in this paper,
are available at \url{https://github.com/EdwardJKim/astroclass}.

\textcolor{blue}{end new text}

@@

We note that \cite{carrascokind2014exhausting} analyzed
different techniques for combining
photometric redshift probability density functions (photo-$z$ PDFs),
and also found that BMC is in general the best
photo-$z$ PDF combination technique.

@@

We thank the referee for a careful reading of the manuscript
and comments that improved this work.
We thank Ignacio Sevilla for helpful and insightful conversations.
We acknowledge support from the National Science Grant No.
\textcolor{red}{XXXX}.
